Hardware-Accelerated Rendering of Intersecting Volumes using Boolean Operations

### 2013-06-13 paper.tex
Right-align numbers in performance table
Fixed some typos

For those reasons, in this paper we examine techniques for rending two volumes
using simple boolean operations.  Although limited, we hope the techniques
presented will be more approachable to software engineers faced with
implementing such technologies, and could help direct future efforts in volume
rendering.

@@

To get the pipeline started, the application feeds vertices to the video card’s
first stage.  So the video card knows which vertices are connected, the
application also provides binding information describing which vertices make up
which polygons.  At this point, the video card’s first stage begins constructing
the polygons and performs some preparatory operations on the vertices, such as
calculating the normal vectors used for illumination and assigning texture
coordinates.  As soon as a polygon is finished, it is immediately passed onto
the next stage, which begins rasterizing the polygons into temporary pixels
called fragments.  The fragments are passed along to the last stage, which
shades the them, decides which ones to discard, and displays those that remain
as pixels.

@@

To apply ray-casting to volume rendering with a high-level programmable shading
language, \cite{Kruger03} adapted the concept to use fragment shaders on a
volume’s bounding box.  Instead of casting rays through a pixel grid, each
fragment generated by rendering the bounding boxes creates its own ray.  The
rays start at the viewpoint and passes through the fragment’s location.  Any
volumes that intersect the ray are sampled at regular intervals along it by
applying the inverse model view matrix to the points and using them as indices
into the volume.  A transfer function converts the values to colors, and each is
blended into the framebuffer at the fragment’s position.  Figure
\ref{fragment-ray-casting} gives a depiction of the process, with green circles
representing fragments.

@@

\begin{table}
  \centering
  \begin{tabular}{ l l l l }
    \toprule
    Approach & Macbook & Desktop & Workstation\\
    \midrule
    Boolean AND with masking & 22 & 13 & 51 \\
    Boolean AND with XOR & 24 & 13 & 76 \\
    Slicing & 6 & 18 & 130 \\
    Optimized Slicing & 25 & 15 & 172 \\
    \bottomrule
  \end{tabular}
  \caption{Performance measurements}
  \label{performance-table}
\end{table}

@@

As shown both Boolean-based approaches outperformed the traditional slicing
method, but were in turn bested by the optimized slicing method.  It should be
noted that the frame rates for the optimized slicing method are the worst case,
and in average it performed even better, around forty frames per second.

>>>>

For those reasons, in this paper we examine techniques for rendering two volumes
using simple boolean operations.  Although limited, we hope the techniques
presented will be more approachable to software engineers faced with
implementing such technologies, and could help direct future efforts in volume
rendering.

@@

To get the pipeline started, the application feeds vertices to the video card’s
first stage.  So the video card knows which vertices are connected, the
application also provides binding information describing which vertices make up
which polygons.  At this point, the video card’s first stage begins constructing
the polygons and performs some preparatory operations on the vertices, such as
calculating the normal vectors used for illumination and assigning texture
coordinates.  As soon as a polygon is finished, it is immediately passed onto
the next stage, which begins rasterizing the polygons into temporary pixels
called fragments.  The fragments are passed along to the last stage, which
shades them, decides which ones to discard, and displays those that remain as
pixels.

@@

To apply ray-casting to volume rendering with a high-level programmable shading
language, \cite{Kruger03} adapted the concept to use fragment shaders on a
volume’s bounding box.  Instead of casting rays through a pixel grid, each
fragment generated by rendering the bounding boxes creates its own ray.  The
rays start at the viewpoint and pass through the fragment’s location.  Any
volumes that intersect the ray are sampled at regular intervals along it by
applying the inverse model view matrix to the points and using them as indices
into the volume.  A transfer function converts the values to colors, and each is
blended into the framebuffer at the fragment’s position.  Figure
\ref{fragment-ray-casting} gives a depiction of the process, with green circles
representing fragments.

@@

\begin{table}
  \centering
  \begin{tabular}{ l r r r }
    \toprule
    Approach & Macbook & Desktop & Workstation\\
    \midrule
    Boolean AND with masking & 22 & 13 & 51 \\
    Boolean AND with XOR & 24 & 13 & 76 \\
    Slicing & 6 & 18 & 130 \\
    Optimized Slicing & 25 & 15 & 172 \\
    \bottomrule
  \end{tabular}
  \caption{Performance measurements}
  \label{performance-table}
\end{table}

@@

As shown both Boolean-based approaches outperformed the traditional slicing
method, but were in turn bested by the optimized slicing method.  It should be
noted that the frame rates for the optimized slicing method are the worst case,
and on average it performed even better, around forty frames per second.
